---
title: "General Social Survey"
author: "Corine Romero"
date: "October 14, 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. Finally address this question.  Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
Both variables in this analysis are categorical; hence we use bar graphs and tables to summarize the data, and a chi-squared test for inference.

$H_0$ : There is not relationship between gun ownership and whether one is in favor of capitol punishment.

$H_0$ : There is a relationship between gun ownership and whether one is in favor of Capitol punishment.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results
```{r}
barchartGC(~CAPPUN+OWNGUN, data=mydata4, main='Capitol Punishment and Gun Ownershipb', type= 'percent')
```

This bar chart shows that those who favor capital punishment and own guns are about equal, but there are more who do not own a gun than oppose Capitol punishment. The bar chart indicates there is a connection between gun ownership and supporting capitol punishment, because those in favor and own guns are about equal.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
tab1= xtabs(formula = ~CAPPUN+OWNGUN, data=mydata4,)
rowPerc(tab1)
```

### Inferential Results
```{r}
chisqtestGC(tab1)
```

The P-value is 0.02 < 0.05 so we reject the null hypothesis and conclude we have evidence to support the connection that gun ownership and whether one favors capitol punishment are related.  The bar chart indicates there is a connection between gun ownership and supporting capitol punishment.

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->


Now answer Is belief in the big bang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
Both variables in this analysis are categorical; hence we use bar graphs and tables to summarize the data, and a chi-squared test for inference.

$H_0$ : There is no relationship between Big bang theory and whether one believes in the afterlife.

$H_0$ : There is a relationship between belief in the big bang theory and the belief in the afterlife.


##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results
```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata4, main='Belief in the Big Bang Theory and Belife in the Afterlife', type= 'percent')
```
The bar chart shows that there is no relationship between in belief in the big bang theory and the belief in afterlife since in both cases there is not a lot of change in the data sets.

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results
```{r}
tab2= xtabs(formula = ~BIGBANG+POSTLIFE, data=mydata4,)
rowPerc(tab2)
```

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results


<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(tab2)
```
The p-value is 0.08 > 0.05 so we fail to reject the null hypothesis and conclude we do not have evidence to support the connection. The bar chart does not support a connection between belief in the big bang theory and belief in the afterlife.

# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Finally determine if a persons opinion on death penalty (CAPPUN) is independent of their race (RACE).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Both variables in this analysis are categorical; hence we use bar graphs and tables to summarize the data, and a chi-squared test for inference.

$H_0$ : There is no relationship between death penalty and race.

$H_0$ : There is a relationship between opinion on the death penalty and race.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results
```{r}
barchartGC(~CAPPUN+RACE, data=mydata4, main='Belief in Capitol Punishment by Race', type= 'percent')
```
The bar chart showed that there is a connection between belief in capitol punishment and race.  since whites favor capitol punishment and blacks oppose capitol punishment.
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
tab3= xtabs(formula = ~CAPPUN+RACE, data=mydata4,)
rowPerc(tab3)
```

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(tab3)
```

Since our p-value is .00009 < 0.05 so we reject the null hypothesis and conclude we have evidence to support the connection between the opinion of capitol punishment and race.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Question #1, Is about the opinion on the death penalty (CAPPUN) is independent of gun ownership (OWNGUN)?  The p-value is 0.02 < 0.05 so we reject the null hypothesis and conclude that we have the evidence to support the connection between gun ownership and capital punishment are related.

Question #2, Is the belief in the big bang theory independent of belief in the afterlife?  The p-value is 0.08 > 0.05 so we have failed to reject the null hypothesis and conclude we do not have enough evidence to support a connection between the belief in the big bang theory and the beige in after life.

Question #3, Is the opinion of capital punishment dependent on race?  Since the p-value is .00009 , 0.05 so we reject the null hypothesis and conclude that we have evidence to support the connection between capital punishment and race.